	\subsection{\sys\ algorithm} 
\label{ssec:ll}

\begin{figure}[!t]
  \centering
  
  \begin{subfigure}[t]{\columnwidth}
      \centering
    \begin{tabular}{|c|c|c|c|c|}
      \hline
      key & value & version & commit& leader\\
      %\hline
      \hline
      k1 & a & 3 & nil &k1\\
      %\hline
      k2 & b & 3 & nil &k1\\
      %\hline
      k3 & c & 3 & nil &k1\\
      \hline
    \end{tabular}
	\caption[]{Tentative transaction}
    \label{fig:model:tentative}
  \end{subfigure}
  
  \begin{subfigure}[t]{\columnwidth}
    \centering
    \begin{tabular}{|c|c|c|c|c|}
      \hline
      key & value & version & commit& leader\\
      \hline
      %\hline
      k1 & a & 3 & 7&k1\\
      %\hline
      k2 & b & 3 & nil &k1\\
      %\hline
      k3 & c & 3 & nil &k1\\
      \hline
    \end{tabular}
	\caption[]{Committed transaction}
    \label{fig:model:committed}
  \end{subfigure}


  \begin{subfigure}[t]{\columnwidth}
    \centering
    \begin{tabular}{|c|c|c|c|c|}
      \hline
      key & value & version & commit& leader\\
      \hline
      %\hline
      k1 & a & 3 & 7&k1\\
      %\hline
      k2 & b & 3 & 7&k1\\
      %\hline
      k3 & c & 3 & 7&k1\\
      \hline
    \end{tabular}
	\caption[]{Post-committed transaction}
    \label{fig:model:postcommit}
  \end{subfigure}

  
  \caption{Evolution of metadata during a transaction. The transaction receives $ts_r=3$ in the begin stage, 
  and the $ts_c$ it receives when committing is $7$; its leader is k1.}
  \label{fig:model}
\end{figure}


\begin{algorithm}[t]
  \begin{algorithmic}
    \begin{small}
      \Procedure{read}{key} 
      \For{rec $\leftarrow$ ds.get(key, versions down from $ts_r$)}
		\If{rec.commit $\not \in\{$nil, abort$\}$} \Comment committed
     			 \If{rec.commit $< ts_r$}  \State return rec.value \EndIf
      		\ElsIf{rec.commit = nil}  
 			\State $ts_c \leftarrow$ {\sc checkLeader}(rec.leader)
   			\State rec.commit $\leftarrow ts_c$ \Comment helping 
 			 \If{$ts_c \not=$ abort  $\wedge\ ts_c <  ts_r$}  \State return  rec.value \EndIf
      		\EndIf
      		


      \EndFor
      \State  return nil
      \EndProcedure

\Statex

      		\Procedure{checkLeader}{leaderPtr} 
      			\State leader $\leftarrow$ ds.get(leaderPtr)
     			 \If{leader $=$ nil} return abort \EndIf
		     	 \If{ leader.commit $\not=$nil} \Comment transaction is complete
		     	 	\State return leader.commit 
		     	 \EndIf
		     	 \Comment try to abort pending transaction
			\State ok $\leftarrow$ ds.check\&mutate(leader.commit, nil, abort)
			\State \If{$\neg$ok} \Comment leader status has changed -- recheck  
				\State return  {\sc checkLeader}({leaderPtr}) \EndIf
			\State return abort
	   	\EndProcedure

\Statex

\Procedure{commit}{}
      
      	\State $ts_c \leftarrow$ TM.commit($ts_r$, write-set) \Comment may return abort
      	\If{$ts_c \not=$ abort}
	      	 \State  ok $\leftarrow$ ds.check\&mutate(leader.commit, nil,  $ts_c$)
   		\If{$\neg$ok} 
   		 $ts_c \leftarrow$ abort \EndIf
	\EndIf
	\State \Comment post-commit
	\ForAll{keys k $\in$ write-set}
			\If{$ts_c =$ abort} ds.remove(\tuple{k, $ts_r$})  	
			\Else\ update  commit field of \tuple{k, $ts_r$} to $ts_c$  in ds%  $\leftarrow ts_c$ 
			\EndIf
	\EndFor
\EndProcedure
      
       \end{small}
  \end{algorithmic}
  \caption{\sys's read and commit operations.} 
  %for transaction with read timestamp $ts_r$.}
  \label{fig:get-pseudocode}
\end{algorithm} 

The meta-data used by \sys\ for transaction management is stored in two additional columns associated with each object version in the data store:
(1) \emph{commit} holds the status of the transaction that wrote this version -- \emph{nil} if it is pending, its  $ts_c$ if committed, and 
otherwise \emph{abort}; 
and
(2) \emph{leader} is a pointer to the writing transaction's commit entry, namely, the \emph{commit} column of the first key written by the same
transaction. 

Since the transaction commit needs to be an atomic step, the first key in the transaction's write-set is chosen to be the leader, and the transaction commits via a single atomic write of $ts_c$ to the leader's commit column. Transactional read operations that find a nil commit entry 
refer to the leader to find out whether the value they are attempting to read has been committed or not.
Figure~\ref{fig:model} shows the metadata of a transaction during its stages of execution. 

\sys's TM is based on Omid's TM, and has the same functionality except that it does not write commit entries.
In particular, it manages a monotonically increasing global clock for allocating read timestamps on begin and commit timestamps on commit,
and it performs validation (i.e., conflict detection) upon commit using an in-memory hash table. 
Since the TM algorithm has been reported elsewhere~\cite{Omid2017}, and the \sys\ client need not be aware of 
its implementation details, we avoid repeating these details here. 

Algorithm~\ref{fig:get-pseudocode} describes \sys's implementation of read and commit operations.
%Begin and write operations follow the template of Algorithm~\ref{alg:schema}. 
The client's operations proceed as follows:

\paragraph{Write.}
The client adds the tentative record \emph{(key, value, $ts_r$, nil , leader)} to the data store, where
the leader is the first key the client updates during the transaction. It also tracks key in its local write-set.


\paragraph{Read.}
The algorithm traverses data  records (using the data store's ds.get API) pertaining
to key with versions smaller than $ts_r$, latest to earliest, and returns the first key that is committed
with a smaller version than $ts_r$. Upon
encountering a tentative record (with commit=nil), the algorithm calls the {\sc checkLeader} function, which
gets the leader's $ts_c$ from the data store, and again returns the key if the $ts_c$ is smaller than $ts_r$. 

If the commit entry is still nil, read cannot return without determining the final commit timestamp
of the pending transaction. 
\Yoni{example here?}. 
To this end, {\sc checkLeader} attempts to forcefully abort the pending transaction. This is done using the
 {check\&mutate} operation, which atomically reads and sets the commit entry to abort in case its value is still nil.
If check\&mutate fails, this means that the transaction is no longer pending, and {\sc checkLeader} is recursively called
to re-read the leader's entry. Note that during the recursive call, the record is no longer nil, and so the recursive call 
returns without attempting another check\&mutate.
 
%When a transaction commits, it atomically checks it hasn't been invalidated before inserting the $ts_c$ to the commit .


\paragraph{Commit.}
The client sends a commit request to the TM, which assigns it a commit timestamp $ts_c$ and checks for conflicts. 
If the TM does not return abort, this means that the validation has been successful, and the client writes $ts_c$ in the leader's {commit} column. 
However, the transaction can commit only if no read has invalidated it. Therefore, it uses 
the data store's check\&mutate function to write $ts_c$ only if the commit status is still nil.
To avoid an extra read of the leader on every read, once a transaction is committed, the post-commit stage writes 
the $ts_c$ to the {commit} columns of all keys in  the transaction's write-set. 
%Following a successful commit, the client adds $ts_c$ to all data items it wrote to.


