%\documentclass[preprint]{sigplanconf-eurosys} 
\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig,endnotes}

%========================
%  Packages
%========================

\usepackage{graphicx,url,color}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{amsthm}  %<---- for a different "look" in theorems (theorem word in bold, etc.) %ACM conflict with proof definition?
%\usepackage{subfigure}
%\usepackage[tight,footnotesize]{subfigure}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{times}

%\usepackage{todonotes}
%\usepackage[normalem]{ulem} %strikethrough: \sout{Hello World}
%\usepackage{lastpage} %for number of pages
%\usepackage{xspace}
%\usepackage{multirow}
%\usepackage{balance}

\usepackage[colorlinks=true,allcolors=blue,breaklinks]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{float}

%========================
%  Macros
%========================

\newcommand{\code}[1]{\textsf{\fontsize{9}{11}\selectfont #1}}

\newcommand{\inred}[1]{{\color{red}{#1}}}
\newcommand{\remove}[1]{}
\newcommand{\Idit}[1]{[[\inred{Idit: #1}]]}
\newcommand{\Yoni}[1]{[\inred{Yoni: #1}]}
\newcommand{\tb}{\hspace{5mm}}

\newcommand{\sys}{Fragola}
\newcommand{\sysll}{OmidLL}
\newcommand{\force}{Force}
\newcommand{\speedup}[1]{#1$\times$}
\newcommand{\tuple}[1]{\ensuremath{\langle \mbox{#1} \rangle}}

%========================
\begin{document}

\date{}

\title{\Large \bf\sys: Low-Latency Transactions in Distributed Data Stores}

\author{
{\rm Yonatan Gottesman\footnotemark[1]\tb Aran Bergman\footnotemark[2]\tb   Edward Bortnikov\footnotemark[1] }\\ 
{\rm Eshcar Hillel\footnotemark[1]\tb Idit Keidar\footnotemark[1] \footnotemark[2]\tb Ohad Shacham\footnotemark[1]}\\
	\footnotemark[1] Yahoo Research\ \ \footnotemark[2] Technion
} % end author


\maketitle



%=========================================================================
%  Abstract
%=========================================================================

\begin{abstract}
We present \sys, a highly scalable transaction processing engine 
for Apache HBase. 
The \sys\ implementation is based on Apache Incubator Omid, but 
improves its protocol to reduce latency 
while at the same time improving throughput.
Under light load, \sys\ is 4x to 5x faster than Omid, and under high loads, it is 
more than an order of magnitude faster. 
 \sys\ further implements a \emph{fast path} for single-key transactions,
processing them almost as fast as native HBase operations, while preserving
transactional semantics relative to longer transactions.

\end{abstract}

%========================

\section{Introduction} \label{sec:intro}
\input{intro1} 

\section{Service API and Semantics} \label{sec:api}
\input{api}

\section{Low-latency Transactions with \sys} \label{sec:ll}
\input{lorra}
\input{implementation}

\section{Fast Path Transactions}
\label{sec:alg}
\input{fast-path}

\section{Evaluation} \label{sec:eval}
\input{Evaluation}

\section{Related Work} \label{sec:related}
\input{related}

\section{Conclusion} \label{sec:conclusions}

As TPSs begin to be used in new application domains, low transaction 
latency becomes an important consideration. 
Motivated by such use cases we developed \sys, a highly scalable 
low-latency transaction processing engine for Apache HBase. 
We implemented \sys\ based on Apache Incubator Omid, and 
improved its protocol to reduce latency (by 4x to 5x under light load
and an order of magnitude under heavy load), 
while also improving throughput. 

We further designed a generic \emph{fast path} for single-key transactions,
which is applicable to other transaction management systems.
Our implementation of the fast path in \sys\ processes single-key
transactions almost as fast as native HBase operations, while preserving
transactional semantics relative to longer transactions.

%\subsection*{Acknowledgments}

\newpage

\bibliographystyle{acm}
\bibliography{main}

\end{document}
