\documentclass[twocolumn,10pt]{article}
%\documentclass[sigconf,10pt]{acmart}
\usepackage{times}
\usepackage{epsfig,endnotes}

\oddsidemargin 4.5pc
\evensidemargin 4.5pc
\advance\oddsidemargin by -1in  % Correct for LaTeX gratuitousness
\advance\evensidemargin by -1in % Correct for LaTeX gratuitousness
\marginparwidth 0pt             % Margin pars are not allowed.
\marginparsep 11pt              % Horizontal space between outer margin and
                                % marginal note

                                % Top of page:
\topmargin 4.5pc                % Nominal distance from top of page to top of
                                % box containing running head.
\advance\topmargin by -1in      % Correct for LaTeX gratuitousness
\headheight 0pt                 % Height of box containing running head.
\headsep 0pt                    % Space between running head and text.
                                % Bottom of page:
\footskip 30pt                  % Distance from baseline of box containing foot
                                % to baseline of last line of text.
\advance\topmargin by -37pt     % Leave 37pt above text for headers
\headheight 12pt                % Height of box containing running head.
\headsep 25pt                   % Space between running head and text.

\textheight 666pt       % 9 1/4 column height
\textwidth 42pc         % Width of text line.
                        % For two-column mode:
\columnsep 2pc          %    Space between columns
\columnseprule 0pt      %    Width of rule between columns.
\hfuzz 1pt              % Allow some variation in column width, otherwise it's
                        % too hard to typeset in narrow columns.



%\setcopyright{none}
%\settopmatter{printacmref=false, printccs=true, printfolios=true}
%\acmDOI{}
% \acmISBN{}
%\acmConference[TechPulse18]{}{}{}
%\renewcommand{\shortauthors}{Shacham et al.}



%========================
%  Packages
%========================

\usepackage{graphicx,url,color}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}  %<---- for a different "look" in theorems (theorem word in bold, etc.) %ACM conflict with proof definition?
%\usepackage{subfigure}
%\usepackage[tight,footnotesize]{subfigure}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[outercaption]{sidecap}    


%\usepackage{todonotes}
%\usepackage[normalem]{ulem} %strikethrough: \sout{Hello World}
%\usepackage{lastpage} %for number of pages
%\usepackage{xspace}
%\usepackage{multirow}
%\usepackage{balance}
%\usepackage[colorlinks=true,allcolors=blue,breaklinks]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
\usepackage{subcaption}
%\usepackage{caption}
%\usepackage{float}
%\usepackage{floatrow}

\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)




%========================
%  Macros
%========================

\newcommand{\code}[1]{\textsf{\fontsize{9}{11}\selectfont #1}}

\newcommand{\inred}[1]{{\color{red}{#1}}}
%\newcommand{\inred}[1]{#1}
\newcommand{\remove}[1]{}
\newcommand{\Idit}[1]{[[\inred{Idit: #1}]]}
\newcommand{\Yoni}[1]{[\inred{Yoni: #1}]}
\newcommand{\tb}{\hspace{10mm}}

\newcommand{\sys}{Omid~FP}
\newcommand{\sysll}{Omid~LL}
\newcommand{\syspc}{Omid~2PC}

\newcommand{\speedup}[1]{#1$\times$}
\newcommand{\tuple}[1]{\ensuremath{\langle \mbox{#1} \rangle}}

\newcommand{\mypara}[1]{ \vspace{8pt} \noindent{\bf #1}\hspace{6pt}}

%========================
\begin{document}

\date{}

\title{Taking Omid to the Clouds: \\ 
Fast,  Scalable Transactions for Real-Time  Analytics }


\author{
Ohad Shacham\ \ 
Yonatan Gottesman\ \ 
Edward Bortnikov\ \ 
Eshcar Hillel\\ 
Yahoo Research, Oath 
\and
Idit Keidar\\
Yahoo Research, Oath and Technion 
} % end author


\maketitle


%=========================================================================
%  Abstract
%=========================================================================

\begin{abstract}


We describe how we evolve Omid, a transaction processing system for Apache HBase, 
to power Apache Phoenix, a cloud-grade real-time SQL analytics engine.  

Omid was originally designed for data processing pipelines at Yahoo, which are, by and large, 
%huge-scale 
throughput-oriented monolithic NoSQL applications. 
%Evolving Omid into 
Providing a platform 
to support 
%a variety of 
converged real-time transaction processing and analytics applications --
dubbed {\em translytics} --  
introduces new functional and performance requirements. For example, SQL 
support is key for developer productivity, multi-tenancy is essential for cloud deployment, 
and latency is cardinal for just-in-time data ingestion and analytics insights.
% in a variety of applications. 

We discuss our efforts to adapt Omid to these new domains, 
as part of 
%which emerged in 
the process of integrating it into Phoenix as the transaction processing backend. A central piece
of our work is latency reduction in Omid's protocol, which also improves  scalability.  
Under light load, the new protocol's latency is 4x to 5x smaller than the legacy Omid's, whereas 
under increased loads it is an order of magnitude faster. 
\end{abstract}

%========================

\section{Introduction} \label{sec:intro}
\input{intro} 

\section{Background} \label{sec:api}
\input{api}

\section{Low-Latency Transactions} \label{sec:ll}
\input{lorra}
\input{implementation}

%\section{Fast-Path Transactions}
%\label{sec:alg}
%\input{fast-path}

%\section{Evaluation} \label{sec:eval}
%\input{evaluation}

\vspace{-0.3cm}
\section{SQL Oriented Features} \label{sec:sql}
\input{sql}

%\section{Related Work} \label{sec:related} \input{related}
\vspace{-0.4cm}
\section{Conclusion} \label{sec:conclusions}

As transaction processing services begin to be used in 
new application domains, low transaction latency and rich SQL semantics
become important considerations. In addition, public cloud deployments 
necessitate solutions compatible with multi-tenancy.
Motivated by such use cases we evolved Omid for cloud use.
As part of this evolution, we
improved Omid's protocol to reduce latency (by up to an order of magnitude) and improve throughput scalability.
%


We have extended Omid with functionalities required in SQL engines, namely 
secondary index construction and multi-snapshot semantics. 
We have successfully integrated it into the Apache Phoenix translytics engine. 

\subsection*{Acknowledgments}
We thank James Taylor for fruitful discussions. 

\bibliographystyle{abbrv}

\bibliography{refs}

\end{document}
