\documentclass[letterpaper,twocolumn,10pt]{article}
%\usepackage{times}
\usepackage{usenix,epsfig,endnotes}

%========================
%  Packages
%========================

\usepackage{graphicx,url,color}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{amsthm}  %<---- for a different "look" in theorems (theorem word in bold, etc.) %ACM conflict with proof definition?
%\usepackage{subfigure}
%\usepackage[tight,footnotesize]{subfigure}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{times}
\usepackage[outercaption]{sidecap}    


%\usepackage{todonotes}
%\usepackage[normalem]{ulem} %strikethrough: \sout{Hello World}
%\usepackage{lastpage} %for number of pages
%\usepackage{xspace}
%\usepackage{multirow}
%\usepackage{balance}

\usepackage[colorlinks=true,allcolors=blue,breaklinks]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
\usepackage{subcaption}
\usepackage{graphicx}
%\usepackage{caption}
\usepackage{float}
%\usepackage{floatrow}


%========================
%  Macros
%========================

\newcommand{\code}[1]{\textsf{\fontsize{9}{11}\selectfont #1}}

\newcommand{\inred}[1]{{\color{red}{#1}}}
\newcommand{\remove}[1]{}
\newcommand{\Idit}[1]{[[\inred{Idit: #1}]]}
\newcommand{\Yoni}[1]{[\inred{Yoni: #1}]}
\newcommand{\tb}{\hspace{5mm}}

\newcommand{\sys}{Fragola}
\newcommand{\sysll}{Omid~LL}
\newcommand{\syspc}{Omid~2PC}

\newcommand{\speedup}[1]{#1$\times$}
\newcommand{\tuple}[1]{\ensuremath{\langle \mbox{#1} \rangle}}

\newcommand{\mypara}[1]{ \vspace{8pt} \noindent{\bf #1}\hspace{6pt}}

%========================
\begin{document}

\date{}

\title{\Large \bf \sys: Low-Latency Transactions in Distributed Data Stores}


\author{
\remove{
{\rm Yonatan Gottesman\footnotemark[1]\tb Aran Bergman\footnotemark[2]\tb   Edward Bortnikov\footnotemark[1] }\\ 
{\rm Eshcar Hillel\footnotemark[1]\tb Idit Keidar\footnotemark[1] \footnotemark[2]\tb Ohad Shacham\footnotemark[1]}\\
	\footnotemark[1] Yahoo Research\ \ \footnotemark[2] Technion  \\ [2mm]
\small Submission Type: Research	
}
} % end author


\maketitle



%=========================================================================
%  Abstract
%=========================================================================

\subsection*{Abstract}

We present \sys, a highly scalable transaction processing engine 
for the Apache HBase key-value store. The \sys\ implementation is based 
on Apache Omid, but improves its protocol to reduce latency 
while at the same time improving throughput. Under light load, 
\sys\ is 4x to 5x faster than Omid. Under increasing loads, it 
is an order of magnitude faster, scaling far beyond Omid's limits.
\sys\ further implements a \emph{fast path} for single-key transactions, 
processing them almost as fast as native HBase operations.



%========================

\section{Introduction} \label{sec:intro}
\input{intro} 

\section{Service API and Semantics} \label{sec:api}
\input{api}

\section{Fast Transactions with \sysll} \label{sec:ll}
\input{lorra}
\input{implementation}

\section{\sys\ Fast Path Transactions}
\label{sec:alg}
\input{fast-path}

\section{Evaluation} \label{sec:eval}
\input{evaluation}

\section{Related Work} \label{sec:related}
\input{related}

\section{Conclusion} \label{sec:conclusions}

As transaction processing services begin to be used in new application domains, low 
transaction latency becomes an important consideration. 
Motivated by such use cases we developed \sys, a highly scalable 
low-latency transaction processing engine for Apache HBase. 
We implemented \sys\ based on Omid, and 
improved its protocol to reduce latency (by 4x to 5x under light load
and an order of magnitude under heavy load). In contrast with 
Omid, it scales almost perfectly on the throughput side. 

We further designed a generic \emph{fast path} for single-key transactions, 
which executes them  almost as fast as native HBase operations (in terms of 
both throughput and latency), while preserving
transactional semantics relative to longer transactions.
Our fast path algorithm is not \sys-specific, and can be similarly supported in 
other transaction management systems.
Our implementation of the fast path in \sys\ can process single-key
transactions  at a virtually unbounded rate (thanks to HBase's horizontal 
scalability), and improves the latency of short transactions
by another 3x--5x under high load.

The \sys\ code is publicly available\footnote{\small{Omitted for blind review.}}. 
%\footnote{\url{https://github.com/yonigottesman/incubator-omid/tree/localTransactions}}, and 
We are working on contributing \sysll\  to the Omid production code base. 
We believe our improvements will become instrumental for modern cloud-based, large-scale, 
latency-sensitive services, e.g., the Phoenix OLTP engine, which are designed 
to harness up to ten thousand nodes.

%\subsection*{Acknowledgments}

\newpage

{\footnotesize \bibliographystyle{acm}

\bibliography{refs}

\end{document}
