\documentclass{vldb}
%\documentclass[letterpaper,twocolumn,10pt]{article}
%\usepackage{times}
\usepackage{epsfig,endnotes}

%========================
%  Packages
%========================

\usepackage{graphicx,url,color}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{amsthm}  %<---- for a different "look" in theorems (theorem word in bold, etc.) %ACM conflict with proof definition?
%\usepackage{subfigure}
%\usepackage[tight,footnotesize]{subfigure}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{times}
\usepackage[outercaption]{sidecap}    


%\usepackage{todonotes}
%\usepackage[normalem]{ulem} %strikethrough: \sout{Hello World}
%\usepackage{lastpage} %for number of pages
%\usepackage{xspace}
%\usepackage{multirow}
%\usepackage{balance}

\usepackage[colorlinks=true,allcolors=blue,breaklinks]{hyperref}   % hyperlinks, including DOIs and URLs in bibliography
\usepackage{subcaption}
%\usepackage{caption}
\usepackage{float}
%\usepackage{floatrow}

\usepackage{balance}  % for  \balance command ON LAST PAGE  (only there!)

% Include information below and uncomment for camera ready
\vldbTitle{Taking Omid to the Clouds: Fast, Scalable Transactions for Real-Time Cloud Analytics}
\vldbAuthors{Shacham et al}
\vldbDOI{https://doi.org/TBD}


%========================
%  Macros
%========================

\newcommand{\code}[1]{\textsf{\fontsize{9}{11}\selectfont #1}}

\newcommand{\inred}[1]{{\color{red}{#1}}}
\newcommand{\remove}[1]{}
\newcommand{\Idit}[1]{[[\inred{Idit: #1}]]}
\newcommand{\Yoni}[1]{[\inred{Yoni: #1}]}
\newcommand{\tb}{\hspace{10mm}}

\newcommand{\sys}{Omid~FP}
\newcommand{\sysll}{Omid~LL}
\newcommand{\syspc}{Omid~2PC}

\newcommand{\speedup}[1]{#1$\times$}
\newcommand{\tuple}[1]{\ensuremath{\langle \mbox{#1} \rangle}}

\newcommand{\mypara}[1]{ \vspace{8pt} \noindent{\bf #1}\hspace{6pt}}

%========================
\begin{document}

\date{}

\title{Taking Omid to the Clouds: \\ 
Fast,  Scalable Transactions for Real-Time Cloud Analytics }

\numberofauthors{6} 
\author{
\alignauthor
Ohad Shacham\\
       \affaddr{Yahoo Research}
       \email{}
\alignauthor
Yonatan Gottesman\\ %\footnotemark[1]
       \affaddr{Yahoo Research}
       \email{}
\alignauthor
Aran Bergman\\
       \affaddr{Technion}
       \email{}
\and
\alignauthor
Edward Bortnikov\\
       \affaddr{Yahoo Research}
       \email{}
\alignauthor
Eshcar Hillel\\
       \affaddr{Yahoo Research}
       \email{}
\alignauthor
Idit Keidar\\
       \affaddr{Technion and Yahoo Research}
       \email{}
} % end author


\maketitle

%=========================================================================
%  Abstract
%=========================================================================

\begin{abstract}

We describe how we evolve Omid, a transaction processing system for Apache HBase, 
to power Apache Phoenix, a cloud-grade real-time SQL analytics engine.  

Omid was originally designed for data processing pipelines at Yahoo, which are, by and large, 
%huge-scale 
throughput-oriented monolithic NoSQL applications. 
%Evolving Omid into 
Providing a platform 
to support 
%a variety of 
converged real-time transaction processing and analytics applications --
dubbed {\em translytics} --  
introduces new functional and performance requirements. For example, SQL 
support is key for developer productivity, multi-tenancy is essential for cloud deployment, 
and latency is cardinal for just-in-time data ingestion and analytics insights.
% in a variety of applications. 

We discuss our efforts to adapt Omid to these new domains, 
as part of 
%which emerged in 
the process of integrating it into Phoenix as the transaction processing backend. A central piece
of our work is latency reduction in Omid's protocol, which also improves  scalability.  
Under light load, the new protocol's latency is 4x to 5x smaller than the legacy Omid's, whereas 
under increased loads it is an order of magnitude faster. We further describe a \emph{fast path} f
protocol or single-key transactions, which enables processing them almost as fast 
as native HBase operations.

\end{abstract}

%========================

\section{Introduction} \label{sec:intro}
\input{intro} 

\section{Background} \label{sec:api}
\input{api}

\section{Low-Latency Transactions} \label{sec:ll}
\input{lorra}
\input{implementation}

\section{Fast-Path Transactions}
\label{sec:alg}
\input{fast-path}

\section{Evaluation} \label{sec:eval}
\input{evaluation}

\section{SQL Oriented Features} \label{sec:sql}
\input{sql}

\section{Related Work} \label{sec:related}
\input{related}

\section{Conclusion} \label{sec:conclusions}

As transaction processing services begin to be used in 
new application domains, low transaction latency and rich SQL semantics
become important considerations. In addition, public cloud deployments 
necessitate solutions compatible with multi-tenancy.
Motivated by such use cases we evolved Omid for cloud use.

As part of this evolution, we
improved Omid's protocol to reduce latency (by 4x to 5x under light load
and an order of magnitude under heavy load) and improve throughput scalability.
%
We further designed a generic \emph{fast path} for single-key transactions, 
which executes them  almost as fast as native HBase operations (in terms of 
both throughput and latency), while preserving
transactional semantics relative to longer transactions.
Our fast path algorithm is not Omid-specific, and can be similarly supported in 
other transaction management systems.
Our implementation of the fast path in \sys\ can process single-key
transactions  at a virtually unbounded rate (thanks to HBase's horizontal 
scalability), and improves the latency of short transactions
by another 3x--5x under high load.

We have extended Omid with functionalities required in SQL engines, namely 
secondary index construction and multi-snapshot semantics. 
We have successfully integrated it into the Apache Phoenix translytics engine. 

\subsection*{Acknowledgments}
We thank James Taylor for fruitful discussions. 

\bibliographystyle{acm}

\bibliography{refs}

\end{document}
